---
title: "Practical Machine Learning - Course Project"
author: "Andreas Hermann"
date: "11/12/2018"
output: html_document
---

## Executive summary

This report shows how a simple machine learning predictor can be built to do human activity recognition based on movement data.
We try to predict the same 5 classes (sitting-down, standing-up, standing, walking, and sitting) with our own model.

We will address these main points:

- Model Building using Cross Validation
- Discussion of Sample Error
- Prediction of Test Cases

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(42)
library(caret)
library(rattle)
library(parallel)
library(doParallel)
```

## Build Model

Load the data and do some exploratory analysis.
```{r}
training <- data.frame(read.csv2('pml-training.csv', sep = ',', dec = '.', na.strings = c('#DIV/0!', 'NA'), stringsAsFactors = TRUE))
testing <- data.frame(read.csv2('pml-testing.csv', sep = ',', dec = '.', na.strings = c('#DIV/0!', 'NA'), stringsAsFactors = TRUE))
```

Browse the data
```{r}
head(training)
str(training)
```

### Data Cleansing

Find the number of incomplete columns
```{r}
length(colnames(training)[colSums(is.na(training)) > 0])
```

Remove columns containing NA values, as it is hard to predict the class for unknown values

```{r}
na.cols <- colnames(training)[apply(training, 2, anyNA)]
training[na.cols] <- NULL
```

Remove some columns that should not be used for the model
```{r}
training['X'] <- NULL
training['cvtd_timestamp'] <- NULL
training['user_name'] <- NULL
training['new_window'] <- NULL
```

Split the data into training and validation sets

```{r}
inTrain <- createDataPartition(training$classe, p=0.8, list=FALSE)
cv_train <- training[inTrain, ]
cv_validation <- training[-inTrain, ]
```

I will train two different types of models. Decision trees and random forests.

## Decision Trees

Train a decision tree with caret package rpart
```{r}
modFitRPart <- train(classe ~ ., data=cv_train, method="rpart")
```

Visualize the trained model

```{r}
fancyRpartPlot(modFitRPart$finalModel)
```


```{r}
pRF <- predict(modFitRPart, newdata=cv_validation)
confusionMatrix(cv_validation$classe, pRF)
```


## Random Forests

Train a Random Forests model using 5-fold cross validation with the random forest method of caret package.
Execute the code in parallel to speed up processing.

```{r}
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)

control <- trainControl(method = "cv", number = 5, allowParallel = TRUE)
modFitRF <- train(classe ~ ., data=cv_train, method="rf", trControl = control)

stopCluster(cluster)
```

Given an accuracy of 99% on the training data for out-of-bag values we should expect a similar accuracy on out of sample data.

Analyse the accuracy of the trained model.

```{r}
pRF <- predict(modFitRF, newdata=cv_validation)
confusionMatrix(cv_validation$classe, pRF)
```


## Sample Error

Accuracy of RPart model was 49% compared to 99% of the RF model.
I select random forest a my model to use for predicting the exercise outcomes.

## Prediction

Predict the class of the test data using our trained random forest model

```{r}
pRF <- predict(modFitRF, newdata=testing)
table(testing$X, pRF)
```
